Longitudinal Study of the Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design

### 2025-03-23 src/main.tex
[text] small changes

\title{%
  % (Very) Preliminary Study of the Longevity of Docker Images from Research Artifacts
  Longitudinal Study of the Software Environment Produced by Dockerfiles from Research Artifacts: Initial Design and Results
}

\author{Quentin Guilloteau}
\email{quentin.guilloteau@inria.fr}
\affiliation{%
  \institution{INRIA}
  \country{France}
}

@@

\subsection{What do we capture?}

@@

\subsection{Regularly Building Containers}

Once the artifacts have been described in the Nickel representation, we are able to build the \df s associated.
We plan to build all the artifacts at the start of each month.
The goal is to build the artifacts for one year.

\todo{building patterns, exponential decay}

With the total computing and energetical costs in mind, we limit the building times of each container to 10 minutes.

@@

@@

For each conference with an artifact evaluation process, we look at all the artifacts and consider only the artifacts using a \df\ to share their software environment.
From the artifact and the \df, we write the associated Nickel file: where is stored the artifact? how to extract the \df\ out of the artifact? which package managers are used?
Once all the Nickel descriptions for the artifacts of a conference have been written, we use \ecg\ to build the download, build, and extract the information out of each of them.
We run \ecg\ the closer to the start of each month as possible.
\ecg\ is executed on the (x86) machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
The local Docker cache of these machine is cleaned before the building of each container.
Each container has an allocated time of 10 minutes to build.

We will execute \ecg\ on each artifact every month for a full year (13 months).

\paragraph{Contributions}

This protocol leaves a lot of room for contributions on the writing of the Nickel descriptions files.

@@

We wrote the description in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a container to share their software environments, and built the associated \df s once every month.
% "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12"
The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, and March 12$^{th}$ 2025.
The deadline to submit the artifacts was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
%\todo{when was the artifact submission, and publicaiton of the papers for europar?}

@@

@@

@@

%\todo{a word on from ubuntu:latest and apt-get update}.

@@

DAPHNE, Antoine?

@@

>>>>

\title{%
  % (Very) Preliminary Study of the Longevity of Docker Images from Research Artifacts
  Longitudinal Study of the Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design% and Results
}

\author{Quentin Guilloteau}
\email{Quentin.Guilloteau@inria.fr}
\affiliation{%
  \institution{INRIA}
  \country{France}
}

@@

\subsection{What do we capture?}\label{sec:ecg:capture}

@@

@@

\subsection{Contributions}

This framework leaves a lot of room for contributions on the writing of the Nickel descriptions files.
The community can contribute by submitting pull requests on the open source repository containing all the Nickel descriptions.

@@

\begin{enumerate}
\item For each conference with an artifact evaluation process, we look at all the artifacts and consider only the artifacts using a \df\ to share their software environment.
\item From the artifact and the \df, we write the associated Nickel file: where is stored the artifact? how to extract the \df\ out of the artifact? which package managers are used?
\item Once all the Nickel descriptions for the artifacts of a conference have been written, we use \ecg\ to build the download, build, and extract the information out of each of them.
\item We run \ecg\ the closer to the start of each month as possible.
\item The local Docker cache of these machine is cleaned before the building of each container.
\item We will execute \ecg\ on each artifact every month for a full year (13 months).
\end{enumerate}

\paragraph{Details}

\begin{itemize}
\item Each container has an allocated time of 10 minutes to build.
\item \ecg\ is executed on the (x86) machines of the \grid\ platform~\cite{grid5000} on the \texttt{dahu} cluster of the Grenoble site.
\end{itemize}

@@

We wrote the description in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and built the associated \df s once every month for six months.
% "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12"
The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, and March 12$^{th}$ 2025.
The deadline to submit the artifacts was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
%\todo{when was the artifact submission, and publicaiton of the papers for europar?}

The following sections follow the dimensions captured by \ecg\ (see Section~\ref{sec:ecg:capture}).

@@

Four out the five articles provided a Zenodo archive to the artifact, the other provided a github link with a specified \texttt{git} tag.

@@

\todo{check which base image, tag of base image and if apt-get update}

@@

@@

Antoine?

@@

This research was funded, in whole or in part, by the European Union’s Horizon 2020 research and innovation programme under grant agreement No. 957407 as DAPHNE.
